---
title: "Exploratory Analysis of loan_data Loans By M Ramadan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.
# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.
library(ggplot2)
library(ggthemes)
library(dplyr)
library(gridExtra)
library(tidyr)
library(MASS)
library(scales)
library(GGally)
```

## Initial Dataset Exploration

```{r echo=FALSE, Load_the_Data}
# Load the Data
getwd()
setwd("D:/Data Anylst Track")
# Read the csv file
loan_data = read.csv("prosperLoanData.csv")
```

```{r echo=FALSE, data_structure}
names(loan_data)
str(loan_data)
nrow(loan_data)
ncol(loan_data)
#factorizing rating as these are the key variable we'd investigate throughout the dataset
loan_data$loan_dataRating.alpha = factor(loan_data$ProsperRating..Alpha.,
                                 levels = c("AA","A","B","C","D","E","HR","NA"))
loan_data$loan_dataRating <-factor(loan_data$ProsperRating..Alpha,
                        levels = c('AA', 'A', 'B', 'C', 'D', 'E', 'HR', 'NA'))
loan_data$loan_dataScore = factor(loan_data$ProsperScore)
```

```{r echo=FALSE, Univariate_Plots}
#Histogram of Prosper Rating by numbers of loans
ggplot(data = na.omit(loan_data), aes(loan_dataRating.alpha)) +
  geom_bar(aes(fill = loan_dataRating.alpha)) +
  ggtitle('Numbers of Loans by Prosper Rating') +
  xlab('Rating') +
  ylab('Number of Loans')
summary(loan_data$loan_dataRating.alpha)
```
 Most of loans have null values , Rating C is most rating exist

```{r echo=FALSE,Prosper_Score_Distribution }
# Check the  Score Distribution
ggplot(data = loan_data, aes(x = ProsperScore)) + 
        geom_bar(color = "black", fill = '#2600e5') +
        theme(axis.text.x = element_text(angle = 60, vjust = 0.6)) +
        xlab("Score") + ggtitle("Loan Score Distribution")
summary(loan_data$ProsperScore)
```
Confirmation of the above, most of the loans are not  rated and rating is  concentrated from 4-8 range.

```{r echo=FALSE, Income_Range}
#Factorized income range 
loan_data$IncomeRange = factor(loan_data$IncomeRange, 
                             levels=c("Not employed", "$0", "$1-24,999", 
                                    "$25,000-49,999", "$50,000-74,999", 
                                    "$75,000-99,999", "$100,000+", 
                                    "Not displayed"))
##3. Income Range
ggplot(data = loan_data, aes(IncomeRange)) +
  geom_bar(color="black", fill = '#2600e5') +
  ggtitle('Borrower Income Range') +
  xlab('Income') +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.6)) +
  ylab('Count')
```
Here we can found that most of Prosper Loan's customers income range is between $25,000-49,999

```{r echo=FALSE, Debt_to_Income_Ratio}
##4. Debt to Income Ratio
ggplot(data = loan_data, aes(x = DebtToIncomeRatio)) +                
        geom_histogram(color = "black", fill = '#007EE5', binwidth = 0.02) +
        xlim(0, quantile(loan_data$DebtToIncomeRatio, prob = 0.99, na.rm=TRUE)) +
        ggtitle("Debt To Income Ratio") +
        xlab("Debt to Income Ratio") +
        ylab("Count")
summary(loan_data$DebtToIncomeRatio)
```
The graph show long-tailed right-skewed. the majority of people have a credit history, and the ratio should be low enough for a secured repayment. 25% of the income is the threshold for most of the borrowers.
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8, Borrower_Purpose}
##5. Borrower's Purpose of Loan 
x <- c('Debt Consolidation', 
                                'Home Improvement','Business', 
                                 'Personal Loan', 
                                 'Student Use', 
                                 'Auto', 
                                 'Baby & Adoption',
                                 'Boat', 
                                 'Cosmetic Procedure', 
                                 'Engagement Ring', 
                                 'Green Loans', 
                                 'Household Expenses', 
                                 'Large Purchases', 
                                 'Medical/Dental', 
                                 'Motorcycle', 'RV',
                                 'Taxes', 'Vacation',
                                 'Wedding Loans', 
                                 'Other', 
                                 'Not Available')
loan_data$ListingCategory <- factor(loan_data$ListingCategory..numeric., 
                                  levels = c(1:6,8:20,7,0), labels = x)
ggplot(loan_data, aes(ListingCategory)) +
  geom_bar(color="black", fill = '#2600e5') +
  ggtitle('The Purpose of the Loan Listing') +
  xlab('Listing Category') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.6)) +
  ylab('Number of loans')
summary(loan_data$ListingCategory)
```

More than 50% of loan are for debt consolidation wich explain the data for income range ,but also we found many people hide the the purpose of loan.

```{r echo=FALSE, message=FALSE,  fig.height=8, warning=FALSE,Loan_by_Year}
# set Loan Orignal Date in Year-Month-Date format
loan_data$LoanOriginationDate <- as.POSIXct(loan_data$LoanOriginationDate,
                                                    format="%Y-%m-%d")
loan_data$LoanOriginationDate.year <-loan_data$LoanOriginationDate %>% 
                                          format('%Y') %>% strtoi()
## 6. Number of Loans split by Year
ggplot(loan_data, aes(as.factor(LoanOriginationDate.year))) +
  geom_bar(color="black", fill = '#2600e5') +
  ggtitle('Number of Loans by Year') +
  xlab('Years') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.6)) +
  ylab('Number of loans')

summary(loan_data$LoanOriginationDate.year)
```

The chart took a plunge in 2009 and exploded the year after and peaked in 2013. Now, this period is coincide with the 2008 Global Financial Crisis and government must have conducted certain action toward credit-typed business. 

```{r echo=FALSE,fig.width=6, fig.height=9, Loan_by_Month}
## 7. Number of Loans split by Month
months <- c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun',
            'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec')

loan_data$LoanOriginationDate.month=format(loan_data$LoanOriginationDate, "%b")

ggplot(loan_data, aes(LoanOriginationDate.month)) +
  geom_bar(color = 'black', fill = '#007EE5') +
  facet_wrap(~LoanOriginationDate.year) +
  ggtitle('Number of Loans by Month') +
  xlab('Month') +
  ylab('Number of Loans') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

very low loan activities in 2009 and the extremely high growth in 2013. Probably in 2013, business went back to normal and the demand for credit increased.

```{r echo=FALSE, Loan_by_Amount}
## 9. Number of Loan Split by Amount
ggplot(loan_data, aes(LoanOriginalAmount)) + 
                geom_histogram(binwidth = 1000) +
                scale_x_continuous(
                limits = c(0,quantile(loan_data$LoanOriginalAmount, 0.99,
                                                        na.rm = TRUE)),
                breaks = seq(0, quantile(loan_data$LoanOriginalAmount, 0.99, 
                                                  na.rm = TRUE), 2000)) +
                theme(axis.text.x = element_text(angle = 90, hjust = 1))
summary(loan_data$LoanOriginalAmount)
```

from the above we found  the minimum loan amoutn is $1,000 USD and people tend to borrow in  4 main ranges ($5000, $10000, $15000, $20000).

```{r echo=FALSE, Employment_status}
##10. Employment status
ggplot(aes(x = EmploymentStatus), data = na.omit(loan_data)) +
                    geom_bar() +
                    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
                    ggtitle("Loans by Employment Status")
summary(loan_data$EmploymentStatus)
```

Among these criteria, probably people would tick in "Employed" box & would not fill-in other details, which also explained why self-employed borrowers are higher than part-time employment.

```{r echo=FALSE, Stated_Monthly_Income}
##11. Stated Monthly Income
ggplot(aes(x = StatedMonthlyIncome), data = loan_data) +
                  geom_histogram(binwidth = 1000) +
                  scale_x_continuous(
                  limits = c(0, quantile(loan_data$StatedMonthlyIncome, 0.99,
                                                              na.rm = TRUE)),
                  breaks = seq(0, quantile(loan_data$StatedMonthlyIncome, 0.99, 
                                                      na.rm = TRUE), 1000)) +
                  theme(axis.text.x = element_text(angle = 90))
summary(loan_data$StatedMonthlyIncome)
```

median income range is at $4667 and mean of $5608 which is quite average for a professional with few years of experience. 

```{r echo=FALSE, Employment_Length_Year}
## 12. Employment length
ggplot(aes(x=EmploymentStatusDuration/12), data=subset(loan_data, 
                    EmploymentStatus %in% 
                    c("Employed", "Full-time", 
                    "Part-time", "Self-employed"))) + 
  geom_histogram(binwidth = 1, fill='#007EE5', color='black') +
  labs(title='Employment Length of Borrower', x='Years of employment') +
  scale_x_continuous(breaks=seq(0, quantile(loan_data$EmploymentStatusDuration, 
                                            0.99, na.rm = TRUE), 1))
summary(loan_data$EmploymentStatusDuration/12)
```

As expected, I see the data is in favor of people have fewer years of employment and as they aging, they do not need much loan as before.

```{r echo=FALSE, Borrower_Rate}
##13. Borrower's Rate
ggplot(data = loan_data, aes(x = BorrowerRate)) + 
        geom_histogram(color = "black", fill = '#007EE5', binwidth = 0.01) +
        xlab("Rate") +
        ggtitle("Borrower Rate") 
summary(loan_data$BorrowerRate)
```

 it's interesting to know the median borrower rate. I'd suspect they have different borrower rate for different term & rating but here we get a median of 18.4%. However, there is a huge notable spike around 31%, which is a very attractive rate for the investors. This is something we can keep in mind while looking for further investigation.

```{r echo=FALSE, Common_Term_Loan}
## 14.  Most common term loan
loan_data$Term <- factor(loan_data$Term)

ggplot(data = loan_data, aes(x = Term)) + 
        geom_bar(fill='#007EE5') +
        ggtitle("Length of the Loan")
summary(loan_data$BorrowerRate)
```

This is an interesting piece of information we would want to investigate further. 36-month term loan is the most sought-after term, if we couple this with the previous common Borrower Rate, a 3 years loan with median 18.4% borrower rate is a very attractive rate for this business.

```{r echo=FALSE, Lender_Yield}
## 15. Investor perspective - Lender Yield
ggplot(aes(x=LenderYield), data=loan_data) + 
    geom_histogram(binwidth=0.005,fill = '#007EE5', color='black') +
    scale_x_continuous(breaks=seq(0, 0.5, 0.05)) + 
    ggtitle("The Lender Yield on the Loan")
```

This chart is quite similar with the previous chat for borrower rate. As someone who's completely new to the industry, I did not fully comprehend there's a very linear and strong relationship between borrower rate and lender yield, which unfold later on during bivariate analysis.

```{r echo=FALSE, Total_Investor}
## 16. Number of investors invest in a loan
ggplot(aes(x=Investors), data= loan_data) + 
    geom_histogram(binwidth=5, fill= '#007EE5', color='black') +
    labs(title="Number of Investors Funded the Loan")
table(loan_data$Investors==1)
```

Out of 113937 loans, about 24.4% are funded individually. This is a very high concentration toward 1 investor. However, if we keep in mind 75% of loan are below $12000. This would make sense.

# Univariate Analysis

#### What is the structure of your dataset?
The dataset is comprised of 88 variables with 113937 observations. The loans cover the period 2005 & 2014. Variables are of classes int, numeric, date, and factor. Although 88 variables seems to be intimidating at first, we can split these variables into 2 main players: the "Borrowers" variables & "Investors" variables.

#### What is/are the main feature(s) of interest in your dataset?
after divde the dataset into two main different aspects :

For Borrower, I believe the Prosper Score, Proser Rating (both numeric & alphabet) are the main indicators of a quality of borrowers. Further bivariate analysis would help to learn about these 2.

For Lender, I now understand Lender Yield is the most important factor for investor.

As a marketplace, Prosper would be more interest in Service Fees & Collection Fees but we would not dive into these variables in our analysis.

#### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
So far, I have seen some evidences that people take more loans in specific months and the interest rate of these loans are strongly related to the term of the each loan. There is something curious about loan with 37%-38% of interest rate and I have not understood why loan with term of 36 month are so much more dispersed than 60 months, although the median of interest rates on 60 month are slightly greater than 36 months.

Also, I didn’t find nothing so interesting about risk score and income level, two variable that I was expecting more. On the other hand, maybe this variables have more to tell when related to others. I would like to understand better why loans with term of 36-months are more dispersed than 60 months. Also I want to find out why risky loans are classified that way.

#### Did you create any new variables from existing variables in the dataset?
yes also I  factorizing a few variables .

#### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
Yes, there is few unsual distribution, I was particularly pay attention to the high spike in lender yield & borrower rate and the spike in LoanOriginalAmount that people tend to buy in bulk. I was factorizing data and since categorical variables enter into statistical models differently than continuous variables, storing data as factors insures that the modeling functions will treat such data correctly.

# Bivariate Plots Section

```{r echo=FALSE, Borrower_Rate_Prospr_Rating}
##19.Prosper Data - BorrowerRate - Prosper Rating
loan_data$ProsperRating..Alpha. <- factor(loan_data$ProsperRating..Alpha.)
ggplot(data = loan_data, aes(x = ProsperRating..Alpha., y = BorrowerRate)) +
        geom_boxplot() +
        xlab("Prosper Rating") +
        ggtitle("Borrower Rate for Different Prosper Rating")
```

As we can see, with better rating and safe borrower, the borrower rate is having an inveser relationship with the prosper rating.

```{r echo=FALSE, Loan_Status_Per_Rating}
#20. Loan Status per Rating
# create a new variable summarizing the result of each loan
loan_data <- loan_data %>% mutate(Status = ifelse(LoanStatus %in%
                     c("Chargedoff", "Defaulted"), 0,
                     ifelse(LoanStatus %in% 
                     c("Completed", "Current", "FinalPaymentInProgress"), 2, 
                     ifelse(LoanStatus %in% 
                     "Cancelled",3,1))))


loan_data$Status <- factor(loan_data$Status, levels = 0:3, 
                         labels = c("Defaulted", 
                                    "Past Due", 
                                    "Current or Paid",
                                    "Cancelled"))

ggplot(data = arrange(loan_data,Status), aes(x = ProsperRating..Alpha., 
                    y = LoanOriginalAmount, fill = Status)) +
                    geom_bar(stat = "identity") +
                    xlab("Prosper Rating") +
                    xlab("Original Loan Amount") +
                    ggtitle("Orignal Loan Amount for Different Prosper Rating")
```

The lower the rank, the more loans get defaulted. The graph might be mistaken as E and HR have less defaulted loans than others, but also keeps in mind loans were less categorized into that area.

```{r echo=FALSE, message=FALSE, warning=FALSE, Yearly_Income_And_Prosper_Rating}
##21. Borrower Profile - Yearly Income ~ Prosper Rating Categorial
ggplot(aes(y = StatedMonthlyIncome *12 , x =ProsperRating..Alpha. ), 
       data = subset(loan_data,loan_data$StatedMonthlyIncome > 0 &
                       loan_data$StatedMonthlyIncome < 9000)) +
      geom_boxplot(aes(fill = ProsperRating..numeric.)) +
      xlab("Prosper Rating") +
      ylab("Yearly Income") +
      ggtitle("Yearly Income for Different Prosper Rating")
summary(loan_data$StatedMonthlyIncome*12)
```

For this graph, we explicitly excluded MonthlyIncome more than 9000 and no income as there're few outliers have monthly income more than 9000 and the chart cannot displayed properly.

```{r echo=FALSE, message=FALSE, warning=FALSE, Employment_Status_Loan_OriginalAmount}
##22. Borrower Profile - Employment Status ~ LoanOriginalAmount
ggplot(aes(x = EmploymentStatus, y = LoanOriginalAmount), data = na.omit(loan_data)) +
                        geom_boxplot() +
                        scale_y_continuous(limits = c(0,15000)) +
                        theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Retired category seems to have higher loan original amount than part-time. This part of data is quite interesting but I don't really know why such event would happen.

```{r echo=FALSE,message=FALSE, warning=FALSE, BorrowerRate_DebtToIncomeRatio}
##23. Borrower Profile - Borrower Rate ~ Debt to Income Ratio
loan_data$DebtToIncomeRatio.bucket <- with(loan_data, 
                                    cut(as.numeric(DebtToIncomeRatio), 
                                         c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 
                                           0.6, 0.7, 0.8, 0.9, 1,
                                           10)))
loan_data$DebtToIncomeRatio.bucket <- factor(loan_data$DebtToIncomeRatio.bucket,
                             labels = c("0-10%","10-20%", "20-30%", "30-40%",
                                      "40-50%", "50-60%", "60-70%", "70-80%",
                                      "80-90%", "90-100%", ">100%"))
ggplot(aes(x = DebtToIncomeRatio.bucket, y = BorrowerRate), data = loan_data) +
                      geom_boxplot() +
                      ylab("Borrower Rate") +
                      xlab ("Debt To Income Ratio Bucket") 
cor(loan_data$DebtToIncomeRatio, loan_data$BorrowerRate, use = 'complete.obs')
```

As much as I want to believe, although there is some sort of relationship between Debt To Income ratio & Borrower Rate, the relationship is not that significant when we do a correlation test. I also did a few back-of-the-envelope graph and shows no significant relationship between BorrowerRate & other factors.

```{r echo=FALSE, message=FALSE, warning=FALSE, LenderYield_Rating}
#24. Investor profile - Lender Yield vs Rating 
loan_data$ProsperRating.alpha = factor(loan_data$ProsperRating..Alpha.,
                                 levels = c("AA","A","B","C","D","E","HR",""))
ggplot(data = na.omit(loan_data), aes(x = ProsperRating.alpha, y = LenderYield)) +
        geom_boxplot() +
                               xlab("Prosper Rating") +
          ggtitle("Yield for Different Prosper Rating")
```

Similar graph with Borrower Rate, nothing worth notice here.

```{r echo=FALSE,message=FALSE, warning=FALSE,LenderYield_ListingCategory}
##25. Investor proflel - Lender Yield vs Listing Category
ggplot(data = loan_data, aes(x = ListingCategory, y = LenderYield)) +
        geom_boxplot() +
        xlab("Category") +
        ylab("Lender Yield") +
        theme(axis.text.x = element_text(angle = 90, vjust = 0.6)) +
        ggtitle("Lender Yield for Different Listing Category")
```

I initially thought there must be better yield for riskier category but it seems like there's no relationship here.


```{r echo=FALSE, message=FALSE, warning=FALSE, LoanOriginalAmount_Term}
##26. Investor Profile - LoanOriginal Amount ~ Term
ggplot(aes(y = LoanOriginalAmount, x = Term), data = loan_data) +
                                            geom_boxplot() +
                                            scale_y_continuous(
    limits = c(0, quantile(loan_data$LoanOriginalAmount, 0.99, na.rm = TRUE)))
```

This chart is a new interesting insight, although the majority of loan are in 36-month term. The Loan original amount is significantly higher for 60 months term. Let's see if it's hold true for Lender Yield

```{r echo=FALSE, LenderYield_Term}
##26. Investor Profile - LenderYield ~ Term
ggplot(aes(y = LenderYield, x = Term), data = loan_data) +
                                            geom_boxplot()
```

Unfortunately, this is not the case. Term Loan have higher Loan Original Amount because it'd need more time to repay those loans.

# Bivariate Analysis

#### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?
I want to explore 2 main variables Lender Yield and Borrower Rate and see which other related variables have the most effect on it. So far, the only relationship I found is through the proprietory Prosper Scoring system. Other factor I was trying to compare was not having any particular relationship.
#### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?
No
#### What was the strongest relationship you found?
The relationship between (Prosper Rating and Lender Yield) and (Prosper Rating and Borrower Rate) has an inversed relationship. The higher the rating, the lower the borrower rate and lender yield.


# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}
#27. Debt To Income Ratio - Prosper Rating - Lender Yield
ggplot(aes(x= DebtToIncomeRatio, y=LenderYield, color=ProsperRating..Alpha.), 
    data=na.omit(filter(loan_data, DebtToIncomeRatio < 1))) +
    geom_point(alpha = 1) +
    #scale_y_log10() +
    #facet_grid(.~ ProsperRating.alpha ) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    ggtitle("Lender Yield  vs Debt to Income Ratio vs Prosper Rate (for loans with rating") +
    xlab ("Debt To Income Ratio < 1") +
    ylab ("Lender Yield") +
    scale_fill_discrete(name = "Prosper Rating")
```

Here is how we take a closer look at Lender Yield vs Prosper Rating and how Prosper Rating was influced by Debt to Income Ratio.

```{r echo=FALSE, fig.width = 12, fig.height = 10,LenderYield_ProsperRate_Term}
#28. Lender Yield vs Prosper Rate vs Term
plot1 <- ggplot(aes(x= ProsperRating..Alpha., y=LenderYield, 
                                  color=ProsperRating..Alpha.), 
       data=na.omit(filter(loan_data, DebtToIncomeRatio < 1))) +
    geom_point(alpha = 0.8, position = "jitter") +
    facet_grid( .~ Term  ) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    ggtitle("Lender Yield vs Term
            and Prosper Rating")
plot2 <- ggplot(aes(x= ProsperRating..Alpha., y= LenderYield ), 
       data=na.omit(filter(loan_data, DebtToIncomeRatio < 1))) +
    geom_boxplot(aes(fill = ProsperRating..Alpha.)) +
    facet_grid( .~ Term   ) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    ggtitle("Boxplot - Lender Yield vs Term and Prosper Rating")
grid.arrange(plot1, plot2, ncol=1, nrow =2)
```

This is a closer look for lender yield vs prosper rating. The majority of loans opt-in for 36-month term and the return for 36-month and 60-month are just higher than 12-month, also considering the fact there're less loan in 12-month term than other term.

```{r echo=FALSE, fig.width = 12, fig.height = 10, BorrowerRate_Year_Term}
#29. Borrower Rate vs Year vs Prosper Rating
plot3 <- ggplot(aes(x= ProsperRating..Alpha., y= BorrowerRate), 
       data=na.omit(filter(loan_data))) +
    geom_boxplot(aes(fill = ProsperRating..Alpha.)) +
    facet_grid( .~ LoanOriginationDate.year) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    ggtitle("Lender Yield vs Term and Prosper Rate")
plot4 <- ggplot(aes(x= ProsperRating..Alpha., y= BorrowerRate, color=ProsperRating..Alpha. ),
    data=na.omit(filter(loan_data))) +
    geom_point(alpha = 1, position = "jitter") +
    facet_grid( .~ LoanOriginationDate.year) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    ggtitle("Lender Yield vs Term and Prosper Rate")
grid.arrange(plot3, plot4, ncol=1, nrow =2)
```

Prosper must have optimized their model throughout the year and as we see the borrower throughout the year, the variation between borrower rate is not that significant anymore and we tend to have smaller standard deviation year-over-year. Something worth noticing is the amount of borrowing suddenly decreased in 2013.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?
Term loan is quite a good indicator whether we have a better Lender Yield or not. Also, we see how three variables Lender Yield, Prosper Rating and Debt To Income Ratio come together and how it affect each order. 
### Were there any interesting or surprising interactions between features?
The criteria for being an AA borrowers seems to be tighten over the year and a there seems to be a fixed borrower rate in criteria HR and AA.
### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.
No
------

# Final Plots and Summary

### Plot One: 

```{r echo=FALSE, fig.width = 12, fig.height = 8, Plot_One}
ggplot(aes(x= DebtToIncomeRatio, y=LenderYield, color=ProsperRating..Alpha.), 
       data=na.omit(filter(loan_data, DebtToIncomeRatio < 1))) +
  geom_point(alpha = 1) +
  #scale_y_log10() +
  #facet_grid(.~ ProsperRating.alpha ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Lender Yield  vs Debt to Income Ratio vs Prosper Rate (for loans with rating)") +
  xlab ("Debt To Income Ratio < 1") +
  ylab ("Lender Yield [%]") +
  scale_fill_discrete(name = "Prosper Rating")
```

### Description One
This basically best summarize the relationship between Lender Yield and Prosper Rating. The higher the risk, the lower the rating, the better the lender yield. We also noticed high rank like AA would not have D/I ration more than 25% and although most borrowers have lower D/I Ratio, there're still high D/I ratio borrowers and fall in lower ProsperRating. Therefore, we see a somewhat upward triangular shape.

### Plot Two

```{r echo=FALSE, fig.width = 12, fig.height = 8,Plot_Two}
ggplot(data = arrange(loan_data,Status), aes(x = ProsperRating..Alpha., 
                    y = LoanOriginalAmount, fill = Status)) +
                    geom_bar(stat = "identity") +
                    xlab("Prosper Rating") +
                    ylab("Original Loan Amount [$]") +
                    ggtitle("Orignal Loan Amount for Different Prosper Rating")
```

### Description Two
This chart confirmed that although there is higher lending yield for lower rating. The investors should also takes a look at the risk of the loan being defaulted. We see very clear in this chart as the rating get worsen, the higher chance the loan getting defaulted.

### Plot Three

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 12, fig.height = 8, Plot_Three}
ggplot(loan_data, aes(as.factor(LoanOriginationDate.year))) +
  geom_bar(color="black", fill = '#2600e5') +
  ggtitle('Number of Loans by Year') +
  xlab('Years') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.6)) +
  ylab('Number of loans')

```

### Description Three

This graph shows a very interesting time in U.S credit business history where most of the banks and financial institution gone under strong investigation from the SEC, only to recover and went back to business and growein stronger than ever.

# Reflection
In this project I selected a dataset from Prosper.com which includes 81 variables and 113,937 observations. With so many variable it was hard to  get started without any financial knowledge. 

The observed default rates in the data set are quite high. Prosper provides ratings for each borrower to indicate the chance that the borrower will repay the complete loan.

Also during the analysis, the impact of the global financial crisis was clear as the data show plunge in 2009 and just start to raise in 2011, and every thin back to normal almost in 2013.

Also found Debt to Income ratio is the predictor of Lending Yield and Borrower Rate and the other variables seems not to play that big of a role. To expand the study, we need to  combine various variables and make a predictive model using logistic regression for a yes or no investment decision or any othe classification .

